Estimating SARS-CoV-2 transmission in England from randomly sampled prevalence surveys

### 2023-12-20 introduction.tex
Fix file names, build, refs

For SARS-CoV-2, measurements of prevalence are noisy because they are based on population sampling.
The sampling is described as the observation process.
The observation process noise is much larger than the noise of the prevalence process (see \cref{inc-prev:sec:observation-process}).
Therefore, the latter is negligible and again two processes need to be considered: infection and observation.

@@

In this chapter, I introduce mechanistic models for estimating the transmission of SARS-CoV-2, an alternative to the statistical approach in \cref{E-backcalc}.
Mechanistic models have a structure which make assumptions on the biological and epidemiological mechanisms behind the infection process (see \cref{E-back-calc:sec:infection-process}), and hence the disease dynamics~\autocite{lesslerMechanistic}.
Individuals are giving the model's state and parameters a biological and/or epidemiological interpretation.
% Of particular interest is that the basic reproduction number, $\R$, and the effective reproduction number, $R_e(t)$, which can both be calculated as functions of the model parameters; using \cref{E-backcalc}'s approach, additional modelling would be required to estimate these numbers.
In particular, this allows changes in transmission to be decomposed into the contributing factors, \eg the number of contacts per day or the probability of transmission upon contact.
The ability to estimate the contribution of each of these factors can inform public health policy.
Furthermore, mechanistic models can be used for scenario-based modelling, which simulates the effect of interventions.
A scenario here is the effect of a proposed intervention (or lack thereof), introduced into the model by modifying the relevant parameters.
The mechanistic model can then be simulated forward in time to understand the intervention's effects.
All of these properties make mechanistic models a useful tool for understanding and controlling infectious diseases.
However, they make stronger assumptions about the disease and population behaviour than statistical models.

The literature on mechanistic models is vast, having been developed for over a century.
The model in \cref{SEIR:sec:SIR} was first formulated by \textcite{kermackContribution}.
However, the earliest mechanistic model to be formulated mathematically was probably \textcite{rossMalariaA}'s model of malaria transmission~\autocite{lesslerMechanistic}.
These early contributions led to important insights for public health.
\Textcite{rossMalariaA} showed that, if mosquitoes were controlled but not eliminated, malaria could be controlled.
\Textcite{kermackContribution} explained the observation that epidemics ended before all individuals were infected due to a population-level build-up of immunity known as \emph{herd immunity}.

@@

I start this chapter by introducing a generic framework for mechanistic modelling of infectious diseases (\cref{SEIR:sec:transmission-generic}).
% This framework is then applied to SARS-CoV-2 in \cref{SEIR:sec:transmission-application}.
It describes the \emph{transmission model}, corresponding to the infection process.
The transmission model needs to be linked to observations; in this chapter, that will be the CIS prevalence data.
The \emph{observation model}, introduced in \cref{SEIR:sec:observation}, is this link; it incorporates both the prevalence process and the observation process.
These two components form the mathematical model, which I then fit to data, requiring specification of an inference process, parameterisation, and priors (\cref{SEIR:sec:inference-implementation}).
This final formulation uses the transmission model from \textcite{birrellRealtime}, but a novel observation model.
I then use a simulation study to check this recovers the parameters successfully (\cref{SEIR:sec:sim-study}).
Having shown that the model can recover the parameters, I then apply it to the CIS data (\cref{SEIR:sec:application}).
Finally, I discuss the results, limitations, and extensions of this work (\cref{SEIR:sec:discussion}).

>>>>

For SARS-CoV-2, measurements of prevalence are noisy because they are based on population sampling.
The sampling is described as the observation process.
The observation process noise is much larger than the noise of the prevalence process (see \cref{E-inc-prev:sec:observation-process}).
Therefore, the latter is negligible and again two processes need to be considered: infection and observation.

@@

In this chapter, I introduce mechanistic models for estimating the transmission of SARS-CoV-2, an alternative to the statistical approach in \cref{E-backcalc}.
Mechanistic models have a structure which make assumptions on the biological and epidemiological mechanisms behind the infection process (see \cref{E-in-prev:sec:infection-process}), and hence the disease dynamics~\autocite{lesslerMechanistic}.
Individuals are giving the model's state and parameters a biological and/or epidemiological interpretation.
% Of particular interest is that the basic reproduction number, $\R$, and the effective reproduction number, $R_e(t)$, which can both be calculated as functions of the model parameters; using \cref{E-backcalc}'s approach, additional modelling would be required to estimate these numbers.
In particular, this allows changes in transmission to be decomposed into the contributing factors, \eg the number of contacts per day or the probability of transmission upon contact.
The ability to estimate the contribution of each of these factors can inform public health policy.
Furthermore, mechanistic models can be used for scenario-based modelling, which simulates the effect of interventions.
A scenario here is the effect of a proposed intervention (or lack thereof), introduced into the model by modifying the relevant parameters.
The mechanistic model can then be simulated forward in time to understand the intervention's effects.
All of these properties make mechanistic models a useful tool for understanding and controlling infectious diseases.
However, they make stronger assumptions about the disease and population behaviour than statistical models.

The literature on mechanistic models is vast, having been developed for over a century.
The model in \cref{E-SEIR:sec:SIR} was first formulated by \textcite{kermackContribution}.
However, the earliest mechanistic model to be formulated mathematically was probably \textcite{rossMalariaA}'s model of malaria transmission~\autocite{lesslerMechanistic}.
These early contributions led to important insights for public health.
\Textcite{rossMalariaA} showed that, if mosquitoes were controlled but not eliminated, malaria could be controlled.
\Textcite{kermackContribution} explained the observation that epidemics ended before all individuals were infected due to a population-level build-up of immunity known as \emph{herd immunity}.

@@

I start this chapter by introducing a generic framework for mechanistic modelling of infectious diseases (\cref{E-SEIR:sec:transmission-generic}).
% This framework is then applied to SARS-CoV-2 in \cref{E-SEIR:sec:transmission-application}.
It describes the \emph{transmission model}, corresponding to the infection process.
The transmission model needs to be linked to observations; in this chapter, that will be the CIS prevalence data.
The \emph{observation model}, introduced in \cref{E-SEIR:sec:observation}, is this link; it incorporates both the prevalence process and the observation process.
These two components form the mathematical model, which I then fit to data, requiring specification of an inference process, parameterisation, and priors (\cref{E-SEIR:sec:inference-implementation}).
This final formulation uses the transmission model from \textcite{birrellRealtime}, but a novel observation model.
I then use a simulation study to check this recovers the parameters successfully (\cref{E-SEIR:sec:sim-study}).
Having shown that the model can recover the parameters, I then apply it to the CIS data (\cref{E-SEIR:sec:application}).
Finally, I discuss the results, limitations, and extensions of this work (\cref{E-SEIR:sec:discussion}).

### 2023-12-20 transmission.tex
Fix file names, build, refs

\begin{document}
\ifSubfilesClassLoaded{
    \setcounter{chapter}{7}
}

@@

This parameterisation takes information on the contact rates, $\kappa_{aa'}(t)$, from several data sources, using pre-existing methodology~\autocites{vanleeuwenTime}{vanleeuwenAugmenting}, which exploits
%; these estimates have been used for previous modelling of the COVID-19 pandemic~\autocites{vanleeuwenTime}{birrellRealtime}.
% These start with estimates of contact in each setting, based on \textcite{vanleeuwenAugmenting}.
pre-pandemic baseline estimates from the POLYMOD study~\autocite{mossongSocial}, and time use, from the UK Time Use Survey (UKTUS)~\autocite{UKTUS}.
POLYMOD asked a representative sample of the UK population about the contacts they make each day, including information on the contacts' age and location (home, work, school, leisure, transport, or other).
UKTUS asked a representative sample of the UK population about how much time they spent doing various activities each day.
\Textcite{vanleeuwenAugmenting} combined these two datasets to estimate the rate of contacts between age groups during each activity.
\Textcite{birrellRealtime} then modify these contact rates based on measured changes in time use during the pandemic.
These changes are measured using Google mobility data~\autocite{googleCOVID19} and school attendance data~(unpublished).
Google mobility data provides the location of Android users over time, relative to a pre-pandemic baseline.
These data sources are then used to down or up-weight the number of contacts in each setting.
Resulting estimates of contacts by setting are then recombined across settings to form $\kappa_{aa'}$
For further details see \textcite{vanleeuwenAugmenting} and \textcite[supplementary material]{birrellRealtime}.

Some behavioural changes (\eg mask wearing) impact transmission.
However, these are not captured by the contact survey.
To allow for these changes as well as uncertainty produced by the procedure above, I follow \textcite{birrellRealtime} in placing a random walk prior on $\log\beta(t)$  on the weekly timescale:
\begin{align}
    \beta(t) &= \beta_{\tilde{w}(t)} \\
    \log\beta_{w+1} &= \log\beta_w + \sigma_\epsilon \epsilon_w \\
    \epsilon_w &\sim \N(0, 1)
\end{align}
where $\tilde{w}(t)$ is the week number of time $t$, with week 1 being the first week considered.
Two hyperparameters will be inferred: $\beta_1$, the initial value of the random walk; and $\sigma_\epsilon$, the standard deviation of the random walk, will have priors placed on them (see \cref{SEIR:sec:priors}).

@@

Also $q_{aa'}$ remains to be specified.
I again follow the approach of \textcite{birrellRealtime}, decomposing it into two independent components such that $q_{aa'} = q^\text{susc}_{a} q^\text{inf}_{a'}$, where
$q^\text{susc}_a$ is the \emph{susceptibility} of stratum $a$, \ie how vulnerable to infection individuals in stratum $a$ are; and
$q^\text{inf}_{a'}$ is the \emph{transmissibility} of stratum $a'$, \ie how likely individuals in stratum $a$ are to pass on an infection.
% Then $\beta^\text{susc}_i$ is the relative risk of an individual in stratum $i$ being infected, relative to an individual in stratum $a_0$.
These measures include both biological differences, \eg due to differences in immune systems, or behavioural, \eg if infants are less hygienic than adults.

@@

Overall the model fits the data well (see \cref{SEIR:fig:prev-young,SEIR:fig:prev-old}).
There are a few features of the data which are not captured by the model.
In particular, prevalence during October is higher in the 16--25s than the model predicts and lower in the 11--16s.
\begin{figure}
    \thisfloatpagestyle{empty}
    \vspace{-3cm}
    \makebox[\textwidth][c]{\includegraphics[width=.9\paperwidth]{SEIR/CIS/prev_young}}
    \captionsetup{width=0.8\paperwidth}
    \captionof{figure}[SEIR prevalence goodness-of-fit (younger ages)]{%
        Goodness-of-fit to prevalence data by region and age for the younger three age groups (see \cref{SEIR:fig:prev-old} for the others).
        The central ribbon and line shows the posterior median and 95\% CrI estimate for the proportion of the relevant strata PCR positive on each day.
        The lighter outer ribbon is the 95\% CrI including the noise due to the sampling (\ie observation process), aggregated by week.
        The dots show the observed prevalence in CIS, aggregated by week.
        For a well-calibrated model, 95\% of the dots would be within the outer ribbon.
    }
    \label{SEIR:fig:prev-young}
\end{figure}
\begin{figure}
    \thisfloatpagestyle{empty}
    \vspace{-3cm}
    \makebox[\textwidth][c]{\includegraphics[width=.9\paperwidth]{SEIR/CIS/prev_old}}
    \captionsetup{width=0.8\paperwidth}
    \captionof{figure}[SEIR prevalence goodness-of-fit (older ages)]{%
        Goodness-of-fit prevalence for the older three age groups (see \cref{SEIR:fig:prev-young} for the other age groups and details).
    }
    \label{SEIR:fig:prev-old}
\end{figure}

@@

The parameters $\beta_t$ absorb all the time-varying changes in transmission not captured by Google mobility and school attendance data.
They trend downwards up until early December when they start increasing again.
Each region is estimated independently, however correlation is clear across the regions.

\begin{figure}
    \includegraphics{SEIR/CIS/beta_walk}
    \caption[Posterior estimates of $\beta_t$]{%
        Posterior estimates $\beta_t$ for each week.
        Shown are ``violins'' where the width is proportional to the kernel-density estimated posterior density at that point.
        Values trend downwards until early December when an increase starts.
    }
    \label{SEIR:fig:beta-walk}
\end{figure}

One MCMC chain for the North West region appeared to get stuck, with very few moves after the first 100,000 iterations.
I reran this chain with a new seed.
Following this, all regions' MCMC chains converged, assessed using the Rhat statistic and ESS.
Rhat for all parameters is < 1.01, except $\sigma_\epsilon$ in London which had a Rhat of 1.018.
All parameters had ESS > 1000, except $\sigma_\epsilon$ in London and the East, which had ESS of 197 and 757 respectively.
As all the parameters of primary interest had high Rhat and ESS, and no parameters had large issues, the inference procedure is considered to have converged.

@@

@@

Unlike most prior work, this modelling is based on a data source that is representative of the general population.
Therefore, the results have fewer biases and informed by data to a greater extent than this work.
There is little-to-no ascertainment bias in the CIS, unlike sources relying on testing behaviour (in the UK, pillar 2 testing).
The CIS allows estimates of transmission in children, unlike sources relying on severe events that occur almost entirely in adults~\autocite{bhopalChildren}.

This approach provides additional understanding, compared to the approach in \cref{E-backcalc}.
For example, decomposing the change in transmission into changing contacts, $\matr{\kappa}(t)$, changing susceptibility $\vec{s}(t)$, and other unmeasured factors, $\beta(t)$.
By understanding which of these components is changing, policies can be better targeted.
If it is shown that $\beta(t)$ decreases, this implies that for a given amount of activity in the population, there is less transmission.
Such decreases might be beneficial as it allows society to operate in a more normal manner without increasing transmission.
Breaking the trade-off between transmission and activity is desirable because it reduces the disease burden without burdesome restrictions on individuals day-to-day lives.

The parameters $\beta_t$ absorb all the time-varying factors not captured by Google mobility and school attendance data.
Many factors contribute to the pattern in the $\beta_t$ values, which have a minima near the start of December.
While this analysis cannot establish causation, several hypotheses can be identified that are worth further investigation.
The decreases in $\beta_t$ over the autumn could be due to increased caution(\eg masking or stricter adherence to social distancing) and/or interventions short of a lockdown as cases and deaths increased~\autocite{jarvisEffect}, or the impact of local measures in the highest transmission areas~\autocite{scottCovid19}.
Local measures are not well captured by $\matr{\kappa}$ because the matrix was for the whole of England.
Allowing each region to have its own $\matr{\kappa}$ may capture the effect of local measures more accurately.
The increases in $\beta_t$ in December coincide with the rise of the Alpha variant~\autocite{lythgoeLineage}, known to be more transmissible~\autocite{daviesEstimated}.
A multiple variant model might be able to capture this rise more accurately.
Such a model would stratify positive tests in the CIS by pre-Alpha or Alpha variant.
Then, each variant would be explicitly included in the model to fit to this data.

These factors are correlated across the country.
This can explain that similar patterns emerge for $\beta_t$ in different regions (see \cref{SEIR:fig:beta-walk}), despite these being estimated independently (\ie there is no model structure which favours correlated values).
These correlations could be exploited to improve the precision of estimates with a more complex model which borrows information across regions.

@@

This discussion makes clear that the results are limited by the number of parameters, currently limited by the inference algorithm.
The algorithm used here tends to struggle with a large number of parameters.
Alternative proposal schemes, such as updating the parameters in blocks\todo{cite what RTM does?}, could help, or a more advanced algorithm such as NUTS (used elsewhere in this thesis).

Extending the analyses carried out here to the rest of the pandemic would clearly be of interest.
I chose the end of the period to coincide with the vaccination rollout.
Vaccination dramatically affects the transmission dynamics by introducing some additional immunity into the population.
Mechanistic models require explicit representation of this immunity\todo{cite something about vaccination}. (Birrell et al, 2023)

@@

>>>>

\begin{document}
\ifSubfilesClassLoaded{
    \setcounter{chapter}{6}
}

@@

This parameterisation takes information on the contact rates, $\kappa_{aa'}(t)$, from several data sources, using pre-existing methodology~\autocites{vanleeuwenTime}{vanleeuwenAugmenting}.
%; these estimates have been used for previous modelling of the COVID-19 pandemic~\autocites{vanleeuwenTime}{birrellRealtime}.
% These start with estimates of contact in each setting, based on \textcite{vanleeuwenAugmenting}.
This method takes pre-pandemic baseline estimates of: $\kappa_{aa'}$ from the POLYMOD study~\autocite{mossongSocial}, and time use from the UK Time Use Survey (UKTUS)~\autocite{UKTUS}.
POLYMOD asked a representative sample of the UK population about the contacts they make each day, including information on each contact's age and location (home, work, school, leisure, transport, or other).
UKTUS asked a representative sample of the UK population about how much time they spent doing various activities (\eg shopping and spending time at home) each day.
\Textcite{vanleeuwenAugmenting} combined these two datasets to estimate the rate of contacts between age groups during each activity.
\Textcite{birrellRealtime} then modify these contact rates based on measured changes in time use during the pandemic.
These changes are measured using Google mobility data~\autocite{googleCOVID19} and school attendance data~(unpublished).
Google mobility data is based on the location of Android users over time, relative to a pre-pandemic baseline.
The mobility and school data are then used to down or up-weight the number of contacts in each setting, and then recombined across settings to form $\kappa_{aa'}$
For further details see \textcite{vanleeuwenAugmenting} and \textcite[supplementary material]{birrellRealtime}.

Some behavioural changes (\eg mask wearing) impact transmission.
However, these are not captured by the contact survey.
To allow for these changes and the uncertainty produced in formulating $\matr{\kaapa}(t)$, I follow \textcite{birrellRealtime} in placing a random walk prior on $\log\beta(t)$  on the weekly timescale:
\begin{align}
    \beta(t) &= \beta_{\tilde{w}(t)} \\
    \log\beta_{w+1} &= \log\beta_w + \sigma_\epsilon \epsilon_w \\
    \epsilon_w &\sim \N(0, 1)
\end{align}
where $\tilde{w}(t)$ is the week number of time $t$, with week 1 being the first week considered.
Two hyperparameters will be inferred: $\beta_1$, the initial value of the random walk; and $\sigma_\epsilon$, the standard deviation of the random walk, will have priors placed on them (see \cref{SEIR:sec:priors}).

@@

Finally,$q_{aa'}$ needs to be specified.
I again follow the approach of \textcite{birrellRealtime}, decomposing it into two independent factors such that $q_{aa'} = q^\text{susc}_{a} q^\text{inf}_{a'}$, where $q^\text{susc}_a$ is the \emph{susceptibility} of stratum $a$, \ie how vulnerable to infection individuals in stratum $a$ are; and $q^\text{inf}_{a'}$ is the \emph{transmissibility} of stratum $a'$, \ie how likely individuals in stratum $a$ are to pass on an infection.
% Then $\beta^\text{susc}_i$ is the relative risk of an individual in stratum $i$ being infected, relative to an individual in stratum $a_0$.
These measures include both biological differences, \eg due to differences in immune systems, or behavioural, \eg if infants are less hygienic than adults.

@@

@@

The parameters $\beta_w$ absorb all the time-varying changes in transmission not captured by Google mobility and school attendance data.
They trend downwards up until early December when they start increasing again.
Each region is estimated independently, however correlation is clear across the regions.

\begin{figure}
    \includegraphics{SEIR/CIS/beta_walk}
    \caption[Posterior estimates of $\beta_w$]{%
        Posterior estimates $\beta_w$ for each week.
        Shown are ``violins'' where the width is proportional to the kernel-density estimated posterior density at that point.
        Values trend downwards until early December when an increase starts.
    }
    \label{SEIR:fig:beta-walk}
\end{figure}

@@

Overall the model fits the data well (see \cref{SEIR:fig:prev-young,SEIR:fig:prev-old}).
There are a few features of the data which are not captured by the model.
In particular, prevalence during October is higher in the 16--25s than the model predicts and lower in the 11--16s.
\begin{figure}
    \thisfloatpagestyle{empty}
    \vspace{-3cm}
    \makebox[\textwidth][c]{\includegraphics[width=.9\paperwidth]{SEIR/CIS/prev_young}}
    \captionsetup{width=0.8\paperwidth}
    \captionof{figure}[SEIR prevalence goodness-of-fit (younger ages)]{%
        Goodness-of-fit to prevalence data by region and age for the younger three age groups (see \cref{SEIR:fig:prev-old} for the others).
        The central ribbon and line shows the posterior median and 95\% CrI estimate for the proportion of the relevant strata PCR positive on each day.
        The lighter outer ribbon is the 95\% posterior predictive interval for the proportion of test results that are positive each week.
        The dots show the observed prevalence in CIS, aggregated by week.
        For a well-calibrated model, 95\% of the dots would be within the outer ribbon.
    }
    \label{SEIR:fig:prev-young}
\end{figure}
\begin{figure}
    \thisfloatpagestyle{empty}
    \vspace{-3cm}
    \makebox[\textwidth][c]{\includegraphics[width=.9\paperwidth]{SEIR/CIS/prev_old}}
    \captionsetup{width=0.8\paperwidth}
    \captionof{figure}[SEIR prevalence goodness-of-fit (older ages)]{%
        Goodness-of-fit prevalence for the older three age groups (see \cref{SEIR:fig:prev-young} for the other age groups and details).
    }
    \label{SEIR:fig:prev-old}
\end{figure}

One MCMC chain for the North West region appeared to get stuck, with very few moves after the first 100,000 iterations (not shown).
I reran this chain with a new seed, with the rerun results presented.
Following this, all regions' MCMC chains converged, assessed using the Rhat statistic and ESS.
Rhat for all parameters is < 1.01, except $\sigma_\epsilon$ in London which had a Rhat of 1.018.
All parameters had ESS > 1000, except $\sigma_\epsilon$ in London and the East, which had ESS of 197 and 757 respectively.
As all the parameters of primary interest had high Rhat and ESS, and no parameters had large issues, the inference procedure is considered to have converged.

@@

% Unlike most prior work (see\todo{ref section with lit review}), this modelling is based on a data source that is representative of the general population.
% % Therefore, the results have fewer biases and informed by data to a greater extent than this work.
% There is little-to-no ascertainment bias in the CIS, unlike sources relying on testing behaviour (in the UK, pillar 2 testing).
% The CIS allows estimates of transmission in children, unlike sources relying on severe events that occur almost entirely in adults~\autocite{bhopalChildren}.

The parameters $\beta_w$ absorb all the time-varying factors not captured by Google mobility and school attendance data.
Many factors contribute to the pattern in the $\beta_w$ values, which have a minima near the start of December.
While this analysis cannot establish causation, several hypotheses can be identified that are worth further investigation.
The decreases in $\beta_w$ over the autumn could be due to increased caution(\eg masking or stricter adherence to social distancing) and/or interventions short of a lockdown as cases and deaths increased~\autocite{jarvisEffect}, or the impact of local measures in the highest transmission areas~\autocite{scottCovid19}.
Local measures are not well captured by $\matr{\kappa}$ because the matrix was for the whole of England.
Allowing each region to have its own $\matr{\kappa}$ may capture the effect of local measures more accurately.
The increases in $\beta_w$ in December coincide with the rise of the Alpha variant~\autocite{lythgoeLineage}, known to be more transmissible~\autocite{daviesEstimated}.
A multiple variant model might be able to capture this rise more accurately.
Such a model would stratify positive tests in the CIS by pre-Alpha or Alpha variant.
Then, each variant would be explicitly included in the model to fit to this data.

These hypotheses for the mechanism causing changes in $\beta_w$ are correlated across the country.
This can explain that similar patterns emerge for $\beta_w$ in different regions (see \cref{SEIR:fig:beta-walk}), despite these being estimated independently (\ie there is no model structure which favours correlated values).
These correlations could be exploited to improve the precision of estimates with a more complex model which borrows information across regions.

@@

This discussion makes clear that the results are limited by the number of parameters, currently limited by the inference algorithm.
\todo{Rephrase this sentence}
The algorithm used here tends to struggle with a large number of parameters.
Alternative proposal schemes, such as updating the parameters in blocks\todo{cite what RTM does?}, could help, or a more advanced algorithm such as NUTS (used elsewhere in this thesis).

Extending the analyses carried out here to the rest of the pandemic would clearly be of interest.
I chose the end of the period to coincide with the vaccination rollout.
Vaccination dramatically affects the transmission dynamics by introducing some additional immunity into the population.
Mechanistic models require explicit representation of this immunity\todo{cite something about vaccination, including new RTM paper}.

@@

Using a mechanistic model provides additional understanding, compared to the approach in \cref{E-backcalc}.
For example, decomposing the change in transmission into changing contacts, $\matr{\kappa}(t)$, changing susceptibility $\vec{s}(t)$, and other unmeasured factors, $\beta(t)$.
By understanding which of these components is changing, policies can be better targeted.
If it is shown that $\beta(t)$ decreases, this implies that for a given contact rate in the population, there is less transmission.
Such decreases might be beneficial as it allows society to operate in a more normal manner without increasing transmission.
Breaking the trade-off between transmission and activity is desirable because it reduces the disease burden without burdesome restrictions on individuals day-to-day lives.
